---
title: "edify_upload"
author: "Candi"
date: "2024-05-08"
output: html_document
---



## Libraries
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(utHelpR)
library(janitor)
library(here)
library(tidyverse)
library(readxl)
library(lubridate)
library(tidyr)
library(dplyr)
library(devtools)
```

## Prior to running all code chunks make sure you are connected to /Share/Research/National Student Clearing House and have saved the returned files in the nsc_return_files directory. Update the file names with the newly returned file names before running the next code chunk.

## Load Return Files
```{r}
nsc_enrolled <- read.csv("/Volumes/Share/Research/National Student Clearing House/nsc_return_files/003671es_598974_DETLRPT_SE_11012024125341_se_loa_nsc_enr_upload.csv")
nsc_applicant <- read.csv("/Volumes/Share/Research/National Student Clearing House/nsc_return_files/003671es_598938_DETLRPT_SE_11012024121858_sa_loa_nsc_app_upload.csv")
nsc_inquiry <- read.csv("/Volumes/Share/Research/National Student Clearing House/nsc_return_files/003671es_598962_DETLRPT_SE_11012024122313_si_loa_nsc_inq_upload.csv")
nsc_prospect <- read.csv("/Volumes/Share/Research/National Student Clearing House/nsc_return_files/003671es_598963_DETLRPT_SE_11012024124936_sp_loa_nsc_pro_upload.csv")
student_id <- utHelpR::get_data_from_sql_file('student_id.sql','edify','project')
```


## Function to standardize return csv files
```{r echo=FALSE, warning=FALSE, message=FALSE}
clean_return_files <- function(input_df){
    output_df <- input_df %>% 
    janitor::clean_names() %>% 
    select(your_unique_identifier, first_name, middle_initial, last_name, name_suffix, requester_return_field, record_found_y_n, search_date, college_code_branch, college_name, college_state, x2_year_4_year, public_private, enrollment_begin, enrollment_end, enrollment_status, class_level, enrollment_major_1, enrollment_cip_1, enrollment_major_2, enrollment_cip_2, graduated, graduation_date, degree_title, degree_major_1, degree_cip_1, degree_major_2, degree_cip_2, degree_major_3, degree_cip_3, degree_major_4, degree_cip_4, college_sequence) %>% 
  mutate(your_unique_identifier = if_else( 
    substr(requester_return_field, 1, 2) == 'BE',
    substr(requester_return_field, 4, 11),
    substr(requester_return_field, 3, 11)))%>% 
  mutate(x2_year_4_year = as.character(x2_year_4_year)) %>% 
  rename(two_year_four_year = x2_year_4_year) %>% 
  mutate(degree_cip_1 = as.character(degree_cip_1)) %>% 
  mutate(degree_cip_2 = as.character(degree_cip_2)) %>% 
  mutate(degree_cip_3 = as.character(degree_cip_3)) %>% 
  mutate(degree_cip_4 = as.character(degree_cip_4)) %>% 
  mutate(nsc_key = paste(requester_return_field, search_date, stringr::str_remove_all(as.character(lubridate::today()), "-"), row_number(), sep = "-"))
  
  return(output_df)
}
```

## Run returned files through function 
```{r}
enrolled <- clean_return_files(nsc_enrolled)
applicant <- clean_return_files(nsc_applicant) 
inquiry <- clean_return_files(nsc_inquiry) 
prospect <- clean_return_files(nsc_prospect)
```

## Union First Two Returned Files 
```{r}
nsc_1 <- union_all(applicant, enrolled)

nsc_2 <- union_all(prospect, inquiry)

nsc_upload <- union_all(nsc_1, nsc_2) %>% 
  left_join(student_id, join_by(your_unique_identifier == prospect_id)) %>% 
  mutate(sis_system_id = coalesce(student_id, your_unique_identifier)) %>% 
  rename(student_id_2 = your_unique_identifier) %>% 
  rename(your_unique_identifier = sis_system_id) %>% 
  relocate(your_unique_identifier, .before = student_id_2) %>% 
  select(!student_id_2) %>% 
  select(!student_id) %>% 
  mutate_at(c('search_date','enrollment_begin','enrollment_end','enrollment_cip_1','enrollment_cip_2','graduation_date','college_sequence'), as.character)
```

## Generate NSC Upload to Edify File
```{r}
file_name <- "nsc_upload.csv"
file_path <- here::here('/Volumes/Share/Research/National Student Clearing House/May 2024/')

write_csv(nsc_upload, paste0(file_path, file_name), append = TRUE, col_names = TRUE , na = '')
```
