---
title: "excel_file_uploads"
author: "Candi Pierce"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(utHelpR)
library(janitor)
library(here)
library(tidyverse)
library(readxl)
library(lubridate)
library(tidyr)
library(dplyr)
```

## Define Cohort from Banner and Slate
### Pull data from file - mutate student_id it join depending on file you may need to change field names
```{r}
de_loa_df <- read.csv(here("sensitive/DE LOA Reinstatement Fall 2025.csv")) %>% 
  mutate(nsc_id = case_when(
    nchar(Student_ID) == 6 ~ paste0("BE000", Student_ID),
    nchar(Student_ID) == 5 ~ paste0("BE0000", Student_ID),
    nchar(Student_ID) == 4 ~ paste0("BE00000", Student_ID),
    TRUE ~ paste0("BE000", Student_ID)  # Default case
  )) %>% 
  rename(last_name = Last_Name, first_name = First_Name, birth_date = BirthDate)
nsc_from_edify_enrolled <- utHelpR::get_data_from_sql_file('nsc_from_edify_enrolled.sql','edify','project')
```

### If file does not have a last term_id or end date select the last completed term
## Double check students with birthdays in the 1900's the conversion may change it to 2000 causing an error in the file.
## Mutate File
```{r}
de_loa_upload_df <- de_loa_df %>% 
  left_join(nsc_from_edify_enrolled, join_by(nsc_id)) %>% 
  mutate(last_name.y = if_else(is.na(last_name.y), last_name.x, last_name.y)) %>% 
  mutate(first_name.y = if_else(is.na(first_name.y), first_name.x, first_name.y)) %>% 
  mutate(birth_date.y = if_else(is.na(birth_date.y), ymd(mdy(birth_date.x)), birth_date.y)) %>% 
  mutate(term_id = if_else(is.na(term_id), '202440', term_id)) %>% 
  mutate(term_end_date = if_else(is.na(term_end_date), as.Date("2024-12-12"), term_end_date)) %>% 
  select(nsc_id, first_name = first_name.y, middle_initial, last_name = last_name.y, name_suffix, 
  birth_date = birth_date.y, term_id, term_end_date)
```



## Declare Variables for NSC Submission File
### Variables
```{r}
school_code <- '003671'
branch_code <- '00'
school_name <- 'Utah Tech University'
query_option <- 'SE'
```

### Record Type and Variables for NSC Header
### Header for submission file
```{r}
header <- tibble(header = 'H1',
                 school_code = school_code,
                 branch_code = branch_code,
                 school_name = str_pad(school_name, width = 20, side = "right"),
                 file_creation_date = str_remove_all(today(), "-"),
                 query_option = query_option,
                 institution = 'I'
                 )
```


## Function to change last names that are "." to NA
```{r}
change_last_name <- function(input_df){
  output_df <- input_df %>% 
  mutate(last_name = if_else(last_name == ".", NA, last_name))
  
  return(output_df)
}
```

## Run files through to update last name from . to NA
```{r}
edify_enrolled <- change_last_name(de_loa_upload_df)
```


## Create Enrolled file for upload
```{r}
enrolled_file <- edify_enrolled %>% 
  mutate(
    detail_records = 'D1',
    ssn = str_pad(' ',9),
    first_name = str_sub(str_pad(str_remove_all(first_name, "[-,.*()`']"), 20, side = 'right'), 1, 20),
    middle_initial = replace_na(substr(str_remove_all(middle_initial, "."), 1, 1), ''),
    last_name =  str_sub(str_pad(replace_na(str_remove_all(last_name, "[-,.*()`']"), NA), 20, side = 'right'), 1, 20),
    suffix = str_pad(replace_na(str_remove_all(name_suffix,"[-,.']"), ''), 5, side = 'right'), 
    birth_date = str_pad(replace_na(str_remove_all(birth_date, "-"),NA), 18 , side = 'left'),
    term_end_date = str_remove_all(term_end_date, "-"),
    blank_column = '',
    school_code = school_code,
    branch_code = branch_code,
    requester_return_field = nsc_id,
    unique_id = duplicated(requester_return_field)
  ) %>%
   select (detail_records, ssn, first_name, middle_initial, last_name, suffix, birth_date, term_end_date, blank_column, school_code, branch_code, requester_return_field)
```


## Drop rows with NA birthdate and last_name.
```{r}
drop_edify_fields <- function(input_df){
  output_df <- input_df %>% 
  drop_na(birth_date) %>% 
  drop_na(last_name) 
 
  return(output_df)
}
```

## Run file through function to drop NA birthdates and last names.
```{r}
enrolled <- drop_edify_fields(enrolled_file)
```



### Record Type and Count for NSC Trailer
### Trailer for enrolled submission file
```{r}
trailer <- tibble(trailer = 'T1',
                  records = nrow(enrolled) + 2
                  )
```

## Generate NSC File
```{r}
file_name <- "se_loa_reinstatement_nsc_upload.tsv"
file_path <- here::here('sensitive/uploads/')

write_tsv(header, paste0(file_path, file_name), col_names = FALSE, na = '')
write_tsv(enrolled, paste0(file_path, file_name), append = TRUE, na = '')
write_tsv(trailer, paste0(file_path, file_name), append = TRUE, na = '')
```

