---
title: "nsc_upload"
author: "Candi"
date: "2024-05-08"
output: html_document
---

## Prior to running all code chunks make sure you are connected to /Share/Research/National Student Clearing House, add a folder with the month and year and update the folder on lines 159, 177, 196, and 215.  These files should be uploaded to the National Student Clearing House in the Fall after October 15th and in the Spring after March 1st.

## Libraries
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(utHelpR)
library(janitor)
library(here)
library(tidyverse)
library(readxl)
library(lubridate)
library(tidyr)
library(dplyr)
```

## Define Cohort from Banner and Slate
### Pull Data from Edify Table nsc_upload
```{r echo=FALSE, warning=FALSE, message=FALSE}
nsc_from_edify_enrolled <- utHelpR::get_data_from_sql_file('nsc_from_edify_enrolled.sql','edify','project')
nsc_from_edify_applicant <- utHelpR::get_data_from_sql_file('edify_applicant.sql','edify','project')
nsc_from_edify_inquiry <- utHelpR::get_data_from_sql_file('nsc_from_edify_inquiry.sql','edify','project')
nsc_from_edify_prospect <- utHelpR::get_data_from_sql_file('nsc_from_edify_prospect.sql','edify','project')
```

## Declare Variables for NSC Submission File
### Variables
```{r}
school_code <- '003671'
branch_code <- '00'
school_name <- 'Utah Tech University'
query_option <- 'SE'
```

### Record Type and Variables for NSC Header
### Header for submission file
```{r}
header <- tibble(header = 'H1',
                 school_code = school_code,
                 branch_code = branch_code,
                 school_name = str_pad(school_name, width = 20, side = "right"),
                 file_creation_date = str_remove_all(today(), "-"),
                 query_option = query_option,
                 institution = 'I'
                 )
```


## Function to change last names that are "." to NA
```{r}
change_last_name <- function(input_df){
  output_df <- input_df %>% 
  mutate(last_name = if_else(last_name == ".", NA, last_name))
  
  return(output_df)
}
```

## Run files through to update last name from . to NA
```{r}
edify_enrolled <- change_last_name(nsc_from_edify_enrolled)
edify_applicant <- change_last_name(nsc_from_edify_applicant)
edify_inquiry <- change_last_name(nsc_from_edify_inquiry)
edify_prospect <- change_last_name(nsc_from_edify_prospect)
```


## Create Enrolled file for upload
```{r}
enrolled_file <- edify_enrolled %>% 
  mutate(
    detail_records = 'D1',
    ssn = str_pad(' ',9),
    first_name = str_sub(str_pad(str_remove_all(first_name, "[-,.*()`']"), 20, side = 'right'), 1, 20),
    middle_initial = replace_na(substr(str_remove_all(middle_initial, "."), 1, 1), ''),
    last_name =  str_sub(str_pad(replace_na(str_remove_all(last_name, "[-,.*()`']"), NA), 20, side = 'right'), 1, 20),
    suffix = str_pad(replace_na(str_remove_all(name_suffix,"[-,.']"), ''), 5, side = 'right'), 
    birth_date = str_pad(replace_na(str_remove_all(birth_date, "-"),NA), 18 , side = 'left'),
    term_end_date = str_remove_all(term_end_date, "-"),
    blank_column = '',
    school_code = school_code,
    branch_code = branch_code,
    requester_return_field = nsc_id,
    unique_id = duplicated(requester_return_field)
  ) %>%
   select (detail_records, ssn, first_name, middle_initial, last_name, suffix, birth_date, term_end_date, blank_column, school_code, branch_code, requester_return_field)
```

## Function to create files for upload
```{r}
create_edify_file <- function(input_df){
  output_df <- input_df %>% 
  mutate(
    detail_records = 'D1',
    ssn = str_pad(' ',9),
    first_name = str_sub(str_pad(str_remove_all(first_name, "[-,.*()`']"), 20, side = 'right'), 1, 20),
    middle_initial = replace_na(substr(str_remove_all(middle_initial, "."), 1, 1), ''),
    last_name =  str_sub(str_pad(replace_na(str_remove_all(last_name, "[-,.*()`']"), NA), 20, side = 'right'), 1, 20),
    suffix = str_pad(replace_na(str_remove_all(name_suffix,"[-,.']"), ''), 5, side = 'right'), 
    birth_date = str_pad(replace_na(str_remove_all(birth_date, "-"),NA), 18 , side = 'left'),
    term_start_date = str_remove_all(term_start_date, "-"),
    blank_column = '',
    school_code = school_code,
    branch_code = branch_code,
    requester_return_field = nsc_id,
    unique_id = duplicated(requester_return_field)
  ) %>%
   select (detail_records, ssn, first_name, middle_initial, last_name, suffix, birth_date, term_start_date, blank_column, school_code, branch_code, requester_return_field)
  
  return(output_df)
}
```

## Run files through function to create files.
```{r}
applicant_file <- create_edify_file(edify_applicant)
inquiry_file <- create_edify_file(edify_inquiry)
prospect_file <- create_edify_file(edify_prospect)
```

## Drop rows with NA birthdate and last_name.
```{r}
drop_edify_fields <- function(input_df){
  output_df <- input_df %>% 
  drop_na(birth_date) %>% 
  drop_na(last_name) 
 
  return(output_df)
}
```

## Run files through function to drop NA birthdates and last names.
```{r}
enrolled <- drop_edify_fields(enrolled_file)
applicant <- drop_edify_fields(applicant_file)
inquiry <- drop_edify_fields(inquiry_file)
prospect <- drop_edify_fields(prospect_file)
```



### Record Type and Count for NSC Enrolled Trailer
### Trailer for enrolled submission file
```{r}
trailer <- tibble(trailer = 'T1',
                  records = nrow(enrolled) + 2
                  )
```

## Generate NSC Enrolled File
```{r}
file_name <- "se_loa_nsc_enr_upload.tsv"
file_path <- here::here('/Volumes/Share/Research/National Student Clearing House/Oct 2024/')

write_tsv(header, paste0(file_path, file_name), col_names = FALSE, na = '')
write_tsv(enrolled, paste0(file_path, file_name), append = TRUE, na = '')
write_tsv(trailer, paste0(file_path, file_name), append = TRUE, na = '')
```


### Record Type and Count for NSC Applicant Trailer
### Trailer for applicant submission file
```{r}
trailer <- tibble(trailer = 'T1',
                  records = nrow(applicant) + 2
                  )
```

## Generate NSC Applicant File
```{r}
file_name <- "sa_loa_nsc_app_upload.tsv"
file_path <- here::here('/Volumes/Share/Research/National Student Clearing House/Oct 2024/')

write_tsv(header, paste0(file_path, file_name), col_names = FALSE, na = '')
write_tsv(applicant, paste0(file_path, file_name), append = TRUE, na = '')
write_tsv(trailer, paste0(file_path, file_name), append = TRUE, na = '')
```


### Record Type and Count for NSC Inquiry Trailer
### Trailer for inquiry submission file
```{r}
trailer <- tibble(trailer = 'T1',
                  records = nrow(inquiry) + 2
                  )
```

## Generate NSC Inquiry File
```{r}
file_name <- "si_loa_nsc_inq_upload.tsv"
file_path <- here::here('/Volumes/Share/Research/National Student Clearing House/Oct 2024/')

write_tsv(header, paste0(file_path, file_name), col_names = FALSE, na = '')
write_tsv(inquiry, paste0(file_path, file_name), append = TRUE, na = '')
write_tsv(trailer, paste0(file_path, file_name), append = TRUE, na = '')
```


### Record Type and Count for NSC ProspectTrailer
### Trailer for prospect submission file
```{r}
trailer <- tibble(trailer = 'T1',
                  records = nrow(prospect) + 2
                  )
```

## Generate NSC Prospect File
```{r}
file_name <- "sp_loa_nsc_pro_upload.tsv"
file_path <- here::here('/Volumes/Share/Research/National Student Clearing House/Oct 2024/')

write_tsv(header, paste0(file_path, file_name), col_names = FALSE, na = '')
write_tsv(prospect, paste0(file_path, file_name), append = TRUE, na = '')
write_tsv(trailer, paste0(file_path, file_name), append = TRUE, na = '')
```

